---
title: "Encelia Data"
---

## Motivation and Context

```{r}
#| label: do this first
#| echo: false
#| message: false

here::i_am("Project-Model-Selection-Encelia.qmd")
```

**Write a few sentences about *Encelia californica*, *Encelia farinosa*, and the Fullerton Arboretum. You may also discuss what you learned about the *iris* dataset.**

*Encelia californica* and *Encelia farinosa* are two bright yellow flowers that can be found in the Fullerton Arboretum. They can be used to restore deserts and native gardens and help with erosion control. It is important that people understand the differences between the two species if they are trying to restore areas and need the correct plant to do so. When restoring areas, it is important that the plant can flourish properly in the environment it is placed in. Both plants are similar in their environments, but there are some vital differences to make sure the plants have the best chances of survival. One of the biggest differences are when they flower. The *Encelia farinosa* flowers from February to May then August to September while *Encelia californica* flowers from February to October. Another difference is the elevations where they thrive. *Encelia farinosa* can go up to 1000 meters above sea level while the *Encelia californica* can go up to 600 meters above sea level.

The Fullerton Arboretum focuses on

## Main Objective

**In your own words, describe the goal of this project.**

## Packages Used In This Analysis

```{r}
#| label: load packages
#| message: false
#| warning: false

library(here)
library(readr)
library(ggplot2)
library(dplyr)
library(rsample)
library(purrr)
library(yardstick)
library(tidyr)
```

| Package | Use |
|------------------------------------|------------------------------------|
| [here](https://github.com/jennybc/here_here) | to easily load and save data |
| [readr](https://readr.tidyverse.org/) | to import the CSV file data |
| [dplyr](https://dplyr.tidyverse.org/) | to massage and summarize data |
| [ggplot2](https://ggplot2.tidyverse.org/) | to create nice-looking and informative graphs |
| [rsample](https://rsample.tidymodels.org/) | to split data into training and test sets |
| [purrr](https://purrr.tidyverse.org) | to run the cross-validation |
| [yardstick](https://yardstick.tidymodels.org) | to evalute the accuracy of the models |
| [tidyr](https://tidyr.tidyverse.org) | to "pivot" the predictions data frame so that each row represents 1 model |

## Design and Data Collection

As a class, we went to the Fullerton Arboretum and identified the two flowers that we wanted to collect the data on. We split into pairs and collected four pieces of data on the two species: the number of rays on each flower, the diameter of the center disk, the diameter of the entire flower, and the stem length. Everything was measured in centimeters. We used a simple measuring ruler and recorded our data on a sheet of paper. The number of rays was counted individually. The diameter of the disk was measured through the center from end to end. The diameter of the entire flower was measured similarly to the disk but extended to include the rays. The stem length was measured from the bottom of the flower part of the plant to the first area that it branched off closest to the flower.

**Briefly describe the limitations of your data collection method. What difficulties did you (or the class collectively) have collecting the data? What choices did you (or the class collectively) make during the data collection process that someone else might choose differently?**

A few limitations of the data collection was limited access. There were many flowers that could not be accessed due to the desire to respect the grounds and not create new and unnatural paths. A lot of the flowers were also close to more dangerous plants like cacti or in areas that were fenced off, so that limited the randomness of the collection. My partner and I decided to stay consistent with who was collecting the data. Since we did this, we tried to limit the variation in measuring judgement that may have occurred had we not stayed consistent, like deciding where the center of a flower was.

## Training-Test Split

**There isn't much data massaging we need to do here, because we were very deliberate about how we set up our data sheet and how we recorded the data on it. However, you may need to do things like import the data and change the type of some variables.**

**At the end of this section, you should write code to randomly split the Encelia data into a training and test set and explain why a training and test set are useful/necessary for this objective.**

```{r}
library(readr)
encelia <- read_csv("~/Documents/Math 237/Data/Encelia Classification Data Collection - Sheet1.csv")
View(encelia)
```

## Exploratory Data Analysis

**Expand on the EDA you did in the Logistic Regression with Encelia activity. Explain what is going on in your graphs and summaries and how those insights relate to the models you will be building in the next section.**

## Modeling

**Propose several logistic regression models and use cross-validation to select a best model, following the steps in the "Cross-Validation (Part 3)" video and the "Model Selection for Logistic Regression" activity.**

E**xplain what your code is doing:**

-   **What is logistic regression? Why are you doing it?**
-   **Why did you choose each model that you are considering?**
-   **Why are you using cross-validation? How does it work?**
-   **Which model are you selecting as the best model? Why?**

**Fit the selected "best" model on the training set and make predictions on the test set. Overall, how accurately is your model making predictions?**

## **Insights**

**Think about how you might visualize the model's predictions and the accuracy of those predictions. Then, create and describe one or more visualizations in line with your ideas.**

**Which flowers (if any) were incorrectly predicted? Why do you suspect that they were incorrectly predicted?**
